---
title: "lab_1"
format: 
  md:
    output: github_document
editor: visual
---


1: Basic Building Blocks

In its simplest form, R can be used as an interactive calculator. Type 5 + 7 and press Enter
```{r}
5+7
```
To assign the result of 5 + 7 to a new variable called x, you type x <- 5 + 7. This can be read as 'x gets 5 plus 7'. Give it a try now.


```{r}
x <- 5 + 7
```
To view the contents of the variable x, just type x and press Enter. Try it now.
```{r}
x
```
Now, store the result of x - 3 in a new variable called y.
```{r}
y <- x - 3
```
What is the value of y? Type y to find out
```{r}
y
```
The easiest way to create a vector is with the c() function, which stands for 'concatenate' or 'combine'. To create a vector containing the numbers 1.1, 9, and 3.14, type c(1.1, 9, 3.14). Try it now and store the result in a variable called z
```{r}
z <- c(1.1, 9, 3.14)
```

Anytime you have questions about a particular function, you can access R's built-in help files via the `?` command. For example, if you want more information on the c() function, type ?c without the parentheses that normally follow a function name. Give it a try.
```{r}
?c
```

Type z to view its contents. Notice that there are no commas separating the values in the output.
```{r}
z
```

You can combine vectors to make a new vector. Create a new vector that contains z, 555, then z again in that order. Don't assign this vector to a new variable, so that we can just see the result immediately.
```{r}
c(z, 555, z)
```

Numeric vectors can be used in arithmetic expressions. Type the following to see what happens: z * 2 + 100.
```{r}
z * 2 + 100
```

Take the square root of z - 1 and assign it to a new variable called my_sqrt.
```{r}
my_sqrt <- sqrt(z - 1)
```

Print the contents of my_sqrt.
```{r}
my_sqrt
```

Now, create a new variable called my_div that gets the value of z divided by my_sqrt.
```{r}
my_div <- z / my_sqrt
```

Go ahead and print the contents of my_div.
```{r}
my_div
```

To see another example of how this vector 'recycling' works, try adding c(1, 2, 3, 4) and c(0, 10). Don't worry about saving the result in a new variable.
```{r}
c(1, 2, 3, 4) + c(0, 10)
```

In many programming environments, the up arrow will cycle through previous commands. Try hitting the up arrow on your keyboard until you get to this command (z * 2 + 100), then change 100 to 1000 and hit Enter. If the up arrow doesn't work for you, just type the corrected command.
```{r}
z * 2 + 1000
```

You can type the first two letters of the variable name, then hit the Tab key (possibly more than once). Most programming environments will provide a list of variables that you've created that begin with 'my'. This is called auto-completion and can be quite handy when you have many variables in your workspace. Give it a try. (If auto-completion doesn't work for you, just type my_div and press Enter.)
```{r}
my_div
```

2: Workspace and Files 

Determine which directory your R session is using as its current working directory using getwd().
```{r}
getwd()
```

List all the objects in your local workspace using ls()
```{r}
ls()
```

Assign 9 to x using x <- 9.
```{r}
x <- 9
```

Now take a look at objects that are in your workspace using ls().
```{r}
ls()
```

List all the files in your working directory using list.files() or dir().
```{r}
dir()
```

As we go through this lesson, you should be examining the help page for each new function. Check out the help page for list.files with the command ?list.files.
```{r}
?list.files
```

Use the args() function to determine the arguments to list.files().
```{r}
args(list.files)
```

Assign the value of the current working directory to a variable called "old.dir".
```{r}
old.dir <- getwd()
```

Use dir.create() to create a directory in the current working directory called "testdir".
```{r}
dir.create("testdir")
```

Set your working directory to "testdir" with the setwd() command.
```{r}
setwd("testdir")
```

Create a file in your working directory called "mytest.R" using the file.create() function.
```{r}
file.create("mytest.R")
```

